\section{This thesis}

This thesis titled \say{Rotational and vibrational action spectroscopic studies on cold molecular ions} discusses action spectroscopic techniques employed to characterise molecular ions in a cryogenic ion trap spectroscopically. Molecular ions relevant to astrochemistry, especially in the interstellar medium and planetary atmospheres, are mainly focused on and are discussed in respective chapters.\\

\textbf{Chapter \ref{chapter:methods}} \emph{\say{Experimental and theoretical methods}}:  gives an overview of the experimental setup used in this study, including the ion source, cryogenic trap and detector. A detailed description of the action spectroscopic methods employed in this thesis to characterize molecular ions' rotational and vibrational transitions is discussed. Technical details such as determining number density with uncertainty, calibration of instruments and instrument setup are discussed in detail.\\

\textbf{Chapter \ref{chapter:C3H3+}} \emph{\say{Laboratory gas-phase vibrational spectra of \texorpdfstring{[C$_3$H$_3$]$^+ $}{[C3H3]+} isomers and isotopologues by IRPD spectroscopy}}: In this chapter, we investigated broadband gas-phase Ne-IRPD spectra of both linear and cyclic forms of [C$_3$H$_3$]$^+$ and reported the first gas-phase IR spectra of the corresponding [C$_3$D$_3$]$^+$ isomers. Various high-level coupled-cluster methods are benchmarked. We also investigated the isomeric ratio quantification of [C$_3$D$_3$]$^+$ produced with different ion source conditions and precursors. 
The results and analysis of this chapter contents are published \cite{Marimuthu2020LaboratorySpectroscopy}.\\

\textbf{Chapter \ref{chapter:CH3CNH+}} \emph{\say{Infrared predissociation  spectroscopy of protonated methyl cyanide,  \texorpdfstring{CH$_3$CNH$^+$ }{CH3CNH+} }}: In this chapter we present a comprehensive experimental and quantum chemical study of the vibrational spectrum of Ne-CH$_3$CNH$^+$. A focus is on the influence of the weakly-bound neon atom on the infrared pre-dissociation experiments. We also demonstrated an efficient computational approach to provide accurate estimates of anharmonic vibrational frequencies of the bare ion and complex. The results and analysis of this chapter contents are published \cite{Marimuthu2021InfraredCH3CNH+}.\\

\textbf{Chapter \ref{chapter:HC3N+}} \emph{\say{A vibrational action spectroscopic study of the Renner-Teller and spin-orbit affected cyanoacetylene radical cation \texorpdfstring{HC$_3$N$^+$}{HC3N+}}}: In this chapter, we present the investigation of the vibrational transitions of HC$_3$N$^+$, an open shell linear molecular ion. The breakdown of the Born-Oppenheimer approximation due to the Renner-Teller (RT) effect (vibrational coupling) is analysed using an effective Hamiltonian approach. The influence of the tag in IRPD, especially on the bending modes of RT-affected open-shell, is discussed in detail. The results and analysis of this chapter contents are published \cite{steenbakkers_vibrational_2023}.\\

In these first chapters, the vibrations transitions on the potential candidates of interstellar molecular ions are 
experimentally and theoretically investigated in detail, and discussed from various perspectives, such as isomer 
quantification, the influence of tag on smaller molecular ions and RT-affected open-shell species. In the following 
chapters, the investigation focuses on high-resolution pure-rotational action spectroscopy. Rotational transitions 
provide distinct molecular fingerprints, and importantly, due to their low excitation temperature, most interstellar 
species are identified through their rotational transitions. The molecular ions discussed above are potential 
interstellar and also (exo-)planetary candidates. On the basis of their vibrational characterization, rotational 
characterization will be followed in the future. Therefore, the following two chapters discuss the implementation and 
investigation of a novel rotational action spectroscopic technique (ROSAA), which utilizes a change in rare-gas atoms 
attachment rates for measuring pure rotational transitions of bare ions. The method is illustrated for a closed and for 
the first time, an open-shell molecular ion. \\

\textbf{Chapter \ref{chapter:CD+}} \emph{\say{Kinetics of CD\texorpdfstring{$^+$}{+} with He buffer gas}}: In this chapter, we report a systematic study and detailed analysis of the CD$^+$ reaction with helium buffer gas atoms with and without the presence of radiation resonant with the $J=0-1$ rotational transition of CD$^+$ ion. This is important in investigating the ROSAA signal intensity (i.e., measured rotational transition intensity) processes using numerical simulation. Consequently, a robust numerical simulation model was developed to predict the intensity of the rotation transition of the molecular ions of interest.\\

\textbf{Chapter \ref{chapter:CO+}} \emph{\say{The Zeeman effect in CO\texorpdfstring{$^+$}{+} observed with rotational action spectroscopy}}: In this chapter, the high-resolution rotational transition of CO$^+$, an open-shell molecular ionic species, is investigated using the ROSAA technique. In addition to an unpaired electron fine structure splitting, a (partly) resolved hyperfine Zeeman splitting is observed due to Earth's magnetic field. The measured signal intensity is investigated using the developed numerical simulation model. The results and analysis of this chapter contents are published \cite{marimuthu_zeeman_2022}.\\
