\section{Theoretical methods}
\label{sec:theory}

In this study, the vibrational and rotational spectroscopic investigations are
supported and complemented by quantum chemical calculations. Therefore, this
section briefly describes the quantum theory of molecular vibration and
rotation using a quantum mechanical approach.\\

In classical mechanics, the system's total energy, which includes kinetic ($T$)
and potential ($V$) energy, is known as the Hamiltonian ($H$), such that

\begin{equation}
    \label{eqn:classical:Hamiltonian}
    H = T + V
\end{equation}

Schr\"odinger postulated the form of Hamiltonian (operator) in quantum
mechanics, as given below

\begin{equation}
    \label{eqn:quantum:Hamiltonian}
    H = - \frac{\hbar^2}{2m} \nabla^2 + V(r)
\end{equation}

where $m$ is mass, $r$ represent the coordinates, $\nabla^2$ is the Laplacian
operate, and the corresponding Schr\"odinger wave equation using $\psi$, the
wavefunction of the quantum state:

\begin{equation}
    \label{eqn:quantum:simple-Hamiltonian}
    H\psi(r, t) = i \hbar \frac{\partial \psi(r, t)}{\partial t}
\end{equation}

One can derive a final time-independent Schr\"odinger equation by \text{substituting}
Eq. \ref{eqn:quantum:Hamiltonian} in Eq. \ref{eqn:quantum:simple-Hamiltonian}

\begin{equation}
    \label{eqn:quantum:wave-eqn}
    H\psi(r) = E\psi(r)
\end{equation}

where $E$ is the eigenvalue of operator $H$, corresponding to the system's
total energy.\\

In quantum mechanics, any molecular state can be fully described by a number of
different degrees of freedom, each of which has a corresponding coordinate
system. In three dimensions, there is the rotational and vibrational motion of
the nuclei and the motion of the electrons. Nuclear and electron spin variables
could also exist.\\

A fundamental idea underlying the description of the quantum states of
molecules is the Born-Oppenheimer (BO) approximation \cite{born_zur_1927},
i.e., the nuclei's ($N$) motion and electrons' motion can be separated since
nuclei are much heavier than the electrons. Under BO approximation, the total
wavefunction ($\psi$) of the molecular system is separable into nuclear
($\psi_n$) and electronic ($\psi_e$) parts, that is

\begin{equation*}
    \psi = \psi_n (R) \psi_{e}(r, R)
\end{equation*}

where $R$ represents an inter-nuclear separation coordinate for each pair of
atoms and $r$ represents the electron's internal coordinates.

The wave function $\psi_N$ can be further factorized into a vibrational
($\psi_v$) and rotational ($\psi_r$) part:

\begin{equation*}
    \psi_n = \psi_v \psi_{r}
\end{equation*}

so that
\[ \psi = \psi_v \psi_{r} \psi_{e} \]

The total energy ($E$) of the system corresponds to the sum of the
contributions from vibrational, rotational and electronic parts, as given
below:

\begin{equation*}
    E =  E_{v} + E_{r} + E_{e}
\end{equation*}

If the molecule possesses net nuclear or electron spin, it is added and
factorised to $E$ and $\psi$, respectively.

However, the BO approximation does not hold true in all cases. For example, the
coupling between electronic and vibrational modes (Renner-Teller effect,
vibronic coupling) is observed for open-shell linear molecular systems (see
chapter \ref{chapter:HC3N+} for treatment of this case).
\input{chapters/technical-details/theory/vibration.tex}
\input{chapters/technical-details/theory/rotation.tex}
\input{chapters/technical-details/theory/calculations.tex}