\chapter{Acknowledgements}
\label{acknowledgements}

I am grateful for the opportunities to work and study in various scientific fields. I am thankful for the support and guidance of those who have helped me along this journey. I want to express my sincere gratitude to all of them.

First and foremost, I would like to express my deepest gratitude to my daily supervisor and co-promoter, dr. Sandra Br\"{u}nken. Her unwavering support, guidance, and encouragement throughout my doctoral journey have left an indelible mark on my academic and professional development. 
Sandra's insightful comments and constructive criticism have been invaluable in completing this thesis. I am incredibly grateful for her continuous support and encouragement.

In addition, I am immensely grateful to my promoter, Prof. dr. Britta Redlich, for her continuous support and insightful feedback on my work. Her encouragement and positive attitude have kept me motivated and on track.

I also want to express my heartfelt appreciation to my parents (Marimuthu and Selvi), brother (Prasanna alias Aarthi), and especially my girlfriend (Sivapriya) for their constant love, understanding, and encouragement throughout my academic journey. Their unwavering faith in my abilities motivated me to persevere through the challenges of completing my Ph.D.

I thank my lab colleagues Dani\"{e}l Rap, Hunarpreet Kaur, Kim Steenbakkers, and Dian Schrauwen at the FIS group at the FELIX laboratory for their assistance and cooperation and for creating a friendly environment overall.

I sincerely thank collaborators, dr. Sven Thorwirth (University of Cologne) and Prof. dr. Wolf Geppert (Stockholm University) for theoretical and experimental support, respectively. And Ph.D. students, especially Franziska Dahlmann (University of Innsbruck), Shreyak Banhatti (University of Cologne) and David Sundelin (Stockholm University), for very friendly collaborating and helping on the part of my research work.

 I want to thank my office colleagues, especially Deepak Pradeep, Maarten Kwaaitaal, Olga Duda and Maxime Gidding, for creating a pleasant and comfortable environment and confabulating on science and other things to keep my every day going. My special thanks go to my previous colleagues, Sjors Bakels, Olga Luschikova and Sandra Wiersma, for their support during the initial stages of my Ph.D.

I must thank the technical staff at FELIX; their help during the night FELIX shifts (Wouter Stumpel, dr. Victor Claessen and Michel Riet) and technical support (Arjan van Vliet and Ren\'{e} van Buuren). And dr. Lex van der Meer and Michel Riet's collaboration for helping in designing and optimizing optical setups simulations. Their willingness to engage in discussions and provide input on technical doubts has dramatically enriched my learning experience.

Finally, I thank everyone who contributed, either directly or indirectly, to the completion of this thesis. My friendships, collaborations, and interactions with many people throughout my Ph.D. journey have profoundly impacted my personal and professional growth. This work would not have been possible without the support and camaraderie of the fantastic people I have been fortunate to encounter during this time.