\section{Rate loss channels}
\label{subsec:rate-loss-channels}
\input{chapters/CD+ kinetics/figs/loss-channel}
The ions are stored up to a maximum of 6 seconds, often this leads to a trap loss of stored ions \cite{mikosch_evaporation_2007, mikosch_evaporation_2008}. To include this loss channel, Eq. \ref{eqn:rate:parent-ion}, \ref{eqn:rate:n-1thcomplexes} and \ref{eqn:rate:nth-complexes} are modified into the following:

\begin{equation}
    \frac{d\cd^+}{dt} = \text{Eq. } \ref{eqn:rate:parent-ion} - R_{loss} \cdot \cd^+
    \label{eqn:rate:modified:parent-ion}
\end{equation}

\begin{equation}
    \frac{d\text{He}_{n-1}\cd^+}{dt} = \text{Eq. } \ref{eqn:rate:n-1thcomplexes} - R_{loss} \cdot \text{He}_{n-1}\cd^+
    \label{eqn:rate:modified:n-1complexes}
\end{equation}

\begin{equation}
    \frac{d\text{He}_{n}\cd^+}{dt} = \text{Eq. } \ref{eqn:rate:nth-complexes} - R_{loss} \cdot \text{He}_{n}\cd^+
    \label{eqn:rate:modified:nth-complexes}
\end{equation}

where $R_{loss}$ indicates the trap loss rate [in \pers] . In addition to the
$R_{loss}$ channel, we have noticed a growth of m/z=30, which has been seen
even at a very low pressure where no higher order complexes except He\CD as
shown in Figure \ref{fig:trap-and-m30-loss-channel-comparision}. Therefore,
this channel is accounted for by adding an additional loss channel, as shown
below:

\begin{equation}
    \frac{d\cd^+}{dt} = \text{Eq. } \ref{eqn:rate:modified:parent-ion} - R_{loss30} \cdot \cd^+
    \label{eqn:rate:parent-ion-mass30-loss}
\end{equation}

The most probable candidate for m/z 30 is DCO$^+$, which could potentially form
at low temperatures by the following reaction with CO$_2$ or H$_2$O neutrals
(as shown for CH$^+$ isotope in KIDA database \cite{wakelam_2014_2015}):

\begin{equation}
    \cd^+ + \text{CO}_2 \xrightarrow{} \text{CO} + \text{DCO}^+
    \label{eqn:mass-30-reaction-with-CO2}
\end{equation}
\begin{equation}
    \cd^+ + \text{H$_2$O} \xrightarrow{} \text{H$_2$} + \text{DCO}^+
    \label{eqn:mass-30-reaction-with-water}
\end{equation}

\begin{equation}
    \frac{d\text{DCO}^+}{dt} = R_{loss30} \cdot \cd^+
    \label{eqn:rate:mass-30-formation}
\end{equation}

Since the m/z 30 also corresponds to the He$_4$\CD complex, Eq.
\ref{eqn:rate:modified:n-1complexes} is modified only when $n-1=4$ (or Eq.
\ref{eqn:rate:modified:nth-complexes} when $n=4$) by adding the corresponding
loss channel \(R_{loss30} \cdot \cd^+\) as shown below:

\begin{equation}
    \frac{d(\text{He}_4\cd^+ + \text{DCO}^+)}{dt} = (\text{Eq. } \ref{eqn:rate:modified:n-1complexes} \text{ or } \ref{eqn:rate:modified:nth-complexes}) + R_{loss30} \cdot \cd^+
    \label{eqn:rate:mass-30-correction}
\end{equation}

The coupled ordinary differential equations for parent ion \CD (Eq.
\ref{eqn:rate:parent-ion}, \ref{eqn:rate:modified:parent-ion} or
\ref{eqn:rate:parent-ion-mass30-loss}) and the formed complexes (Eq.
\ref{eqn:rate:n-1thcomplexes}, \ref{eqn:rate:nth-complexes} or
\ref{eqn:rate:mass-30-correction}) are used to numerically simulate and fit the
experimentally observed data to determine the rate constants, which is
discussed in more detail in the next section in different aspects. This process
is vital for understanding the ROSAA spectroscopic technique employed in this
work.